Prediction of PIK3CA mutation with gene expression

### 2020-05-20 plosOne/S1Appendix.Rmd
author edit

---
title: "Supplimentary"
author: "JK"
date: '2020 4 3 '
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

Data repository

Pancancer TCGA RNA data: http://api.gdc.cancer.gov/data/3586c0da-64d0-4b74-a449-5ff4d9136611

>>>>

---
title: "Supplimentary"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

### 2020-05-20 plosOne/plosOne.Rmd
author edit

---
title: Prediction of PIK3CA mutation with gene expression
author:
  - name: Jun Kang
    email: jkang.alien@gmail.com
    affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital, College of Medicine, The Catholic University of Korea, Seoul, South Korea
    corresponding: jkang.alien@gmail.com
abstract: |
bibliography: [mybibfile.bib, pik3ca.bib, rpackage.bib, genes.bib] 
indent: true
output: rticles::plos_article
csl: plos.csl
---

@@

- Brevity
- Logic and clarity
- Clean typing

- The problem  
  * PIK3CA mutation in selecting drug

@@

- The main message answers the question and main supporting evidence

@@

- Critical assessment opinions on any shortcomings in study design

@@

- limitations in methods

  * Case imbalance  
  * flaws in analysis  
  * validity of assumption

- Comparison with other studies where inconsistencies are discussed  
- Evaluate the results - not the authors

Our PIK3CA prediction model performed similar with the RAS activation prediction model of the previous study in terms of AUROC (0.84). However AUPR of our model for PIK3CA is lower than model for RAS activation (0.39 vs 0.63). The reason of lower AUPR of our model can be explained by low prevalence rate of PIK3CA mutation and imbalanced dataset. The model for RAS activation trained with cancer types with more than 0.05 prevalence of RAS activation to avoid imbalance classification problem. We included all cancer types into modeling process. The lower prevalance rate of target variable mean our dataset has a lower baseline of AUPR. In the subanalysis for performance of each cancer types, the cancer types with higher PIK3CA mutation rate showe better AUPR. The imbalanced data of our model

@@

- Conclusions comments on possible biological or clinical implications and suggestions for further research.

@@

- S1 Appendix.  
- S2 Figure.  
- S1 Table.  
- S2 Table.  
- S3 Table.

>>>>

---
title: Prediction of PIK3CA mutation with gene expression
author:
  - name: Jun Kang
    email: jkang.alien@gmail.com
    affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital, College of Medicine, The Catholic University of Korea, Seoul, South Korea

  - name: Ahwon Lee
    email: klee@catholic.ac.kr
    affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital, College of Medicine, The Catholic University of Korea, Seoul, South Korea

  - name: Youn Soo Lee
    email: lys9908@catholic.ac.kr
    affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital, College of Medicine, The Catholic University of Korea, Seoul, South Korea
    corresponding: lys9908@catholic.ac.kr

abstract: | 
  Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor-positive HER2 negative subtype. We apply a supervised elastic net penalized logistic regression model in prediction PIK3CA mutation from gene expression data. We used the TCGA pan-cancer dataset. Penalized logistic regression was applied to prediction modeling. 10-fold cross-validation with targe variable stratification was done over the hyperparameter grid. 10845 cases were available for both PIK3CA mutation and mRNA expression data. In 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (Ridge regression) showed best performance in terms of AUROC. The final model was trained with the selected hyperparameters with the entire trainset. The trainset AUROC was 0.93 and the testset AUROC was 0.84. The AUPR of trainset was 0.66 and the testset AUPR was 0.39. Both IGFR1R and PTEN are key players in the tyrosine kinase pathway. The cancer type was an important predictor. Our study suggests that the prediction of genomic alteration with gene expression data is possible with good performance.

bibliography: [mybibfile.bib, pik3ca.bib, rpackage.bib, genes.bib] 
indent: true
output: rticles::plos_article
csl: plos.csl
---

@@

@@

@@

@@

Our PIK3CA prediction model performed similar with the RAS activation prediction model of the previous study in terms of AUROC (0.84). However AUPR of our model for PIK3CA is lower than model for RAS activation (0.39 vs 0.63). The reason of lower AUPR of our model can be explained by low prevalence rate of PIK3CA mutation and imbalanced dataset. The model for RAS activation trained with cancer types with more than 0.05 prevalence of RAS activation to avoid imbalance classification problem. We included all cancer types into modeling process. The lower prevalance rate of target variable mean our dataset has a lower baseline of AUPR. In the subanalysis for performance of each cancer types, the cancer types with higher PIK3CA mutation rate showe better AUPR.

@@

@@

- S1 Appendix.

### 2020-05-20 plosOne/plosOne.tex
author edit

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Prediction of PIK3CA mutation with gene expression} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Jun Kang\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital, College of
Medicine, The Catholic University of Korea, Seoul, South Korea}}\textsuperscript{*}\\
\bigskip
\bigskip
* Corresponding author: jkang.alien@gmail.com\\
\end{flushleft}
% Please keep the abstract below 300 words

@@

\begin{itemize}
\item
  Brevity
\item
  Logic and clarity
\item
  Clean typing
\item
  The problem

  \begin{itemize}
  \tightlist
  \item
    PIK3CA mutation in selecting drug
  \end{itemize}
\end{itemize}

@@

The machine learning approach has been actively researched to detect
genomic alterations. Machine learning can build a prediction model from
a large number of predictors such as radiolomic features {[}1{]},
pathology image {[}2{]} or gene expression data {[}3{]}. Because the
most direct genomic test is more specific and sensitive than predictive
models, the machine learning approach might have a limited role in
clinical practice. However, the machine learning prediction can be a
second-best when the direct test fails.

@@

We used the TCGA pan-cancer dataset. TCGA is a cancer genomic consortium
that archives data of exome sequencing, gene expression, DNA
methylation, protein expression, and clinical data of more than 10000
cancer samples across 33 common cancer types. The gene expression TCGA
pan-cancer dataset is batch-corrected with normalization. The TCGA
dataset is publically available. PIK3CA mutation data was got using
cgdsr rpackage.{[}8{]} Gene expression data was got from the GDAC
firehose using the RTCGAToolbox R package. {[}9{]}

10845 cases were available for both PIK3CA mutation and mRNA expression
data. 5128 out of 20502 genes were included in the modeling process
after filtering with median absolute deviation as described at the
modeling process method. 33 cancer type dummy variables were included in
predictor variables.

@@

To narrow down potential predictors, Genes with a large the median
absolute deviation (more than third-quartiles) were selected.
Yeo-Johnson transformation was done to correct skewness. Centering and
scaling were done. All preprocessing was done using recipe r package.
{[}10{]} Penalized logistic regression was applied to prediction
modeling. 10-fold cross-validation with targe variable stratification
was done over the hyperparameter grid: \(\lambda\) \{\(10^{-5}\),
\(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
\(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Lambda is penalty scaling parameter
and alpha is mixing parameter of penalty function
(\((1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1\)).
{[}11{]}

@@

Model performance was evaluated with the area under the receiver
operating characteristic (ROC) and area under the ROC (AUROC) and
precision-recall (PR) and area under the curve (AUPR). The prevalence of
PIK3CA mutation is low. The low prevalence of PIK3CA results in an
imbalanced dataset which makes the prediction difficult. The AUROC is
optimistic in terms of performance. The AUPR is more informative than
AUROC on imbalanced datasets. {[}12{]} The modeling process was done
with tidymodels rpackage. {[}13{]}

@@

\hypertarget{prevalence-rate-of-pik3ca-mutation}{%
\subsection{prevalence rate of PIK3CA
mutation}\label{prevalence-rate-of-pik3ca-mutation}}

The prevalence rate of PIK3CA was 0.11 in all cases. The PIK3CA
prevalence rate of each cancer type varied. The median prevalence rate
of PIK3CA of each cancer type was 0.03 (range 0-0.33) (Figure 1).

@@

In 10-fold cross-validation, the model with \(\lambda\) = 0.01 and
\(\alpha\) = 1.0 (Ridge regression) showed best performance in terms of
AUROC. The final model was trained with the selected hyperparameters
with the entire trainset.

The trainset AUROC was 0.93 and the testset AUROC was 0.84. The AUPR of
trainset was 0.66 and the testset AUPR was 0.39. (Figure 1A)

@@

Because the prevalence of PIK3CA mutation varies across the cancer type,
the performance of each cancer type was investigated. The AUROC and AUPR
were positively correlated between the train set and test set in cancer
type subanalysis. (Figure 1B) The AUPR was high in cancer type with high
PIK3CA mutation rate such as colon, breast, Uterus cancer types. The
AUROC did not correlate with the PIK3CA mutation rate of each cancer
type. (Figure 1C)

@@

Figure 2 shows the top 30 important predictors. The coefficient is the
parameter of the predictor which represents the effect of the predictor
on prediction. IGF1R mRNA expression was the strongest negative
predictor and PTEN was the strongest positive predictor. Both genes and
PIK3CA are key players in the tyrosine kinase pathway. The cancer type
was an important predictor. Some cancer types including uterine
carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA), pancreatic
adenocarcinoma (PAAD), lymphoid neoplasm diffuse large B-cell lymphoma
(DLBC) are the strongest predictors.

@@

\begin{itemize}
\tightlist
\item
  The main message answers the question and main supporting evidence
\end{itemize}

Our model showed good performance to predict the PIK3CA mutation of
various cancer types. This result shows that the supervised elastic net
penalized logistic regression model can be applied not only to the RAS
activation pathway but also PIK3CA mutation. Both the RAS activation
pathway and PIK3CA mutation are important and common cancer genomic
alterations. They have a significant effect on gene expression in cancer
cells. It might be a challenging prediction of genomic alterations that
are infrequent or have a weak effect on gene expression.

\begin{itemize}
\tightlist
\item
  Critical assessment opinions on any shortcomings in study design
\end{itemize}

Prediction modeling from the TCGA pan-cancer dataset has limitations
regarding data preprocessing. Methods for cancer gene expression have
been developed for research. To statistical analysis, the gene
expression data is processed between-sample normalization to remove the
batch effect. If the model has been trained from the between-sample
normalization, a new sample can not be exactly processed like trainset.
A model based on gene expression TCGA pan-cancer dataset has a
limitation on preprocessing. It is necessary for developing a
preprocessing method that is independent with a dataset to apply the
gene expression data to the prediction model.

\begin{itemize}
\item
  limitations in methods

  \begin{itemize}
  \tightlist
  \item
    Case imbalance\\
  \item
    flaws in analysis\\
  \item
    validity of assumption
  \end{itemize}
\item
  Comparison with other studies where inconsistencies are discussed\\
\item
  Evaluate the results - not the authors
\end{itemize}

Our PIK3CA prediction model performed better than the RAS activation
prediction model of the previous study in terms of both AUROC (0.84 vs
0.75) and AUPR (0.39 vs 0.24) on test set prediction. Our test set is
corresponding to the samples initially filtered from training. The
target variable of our study is more specific than the previous study.
The specific important mutations can affect stronger downstream gene
expression than the broad events pathway activation. The previous study
might be a more difficult prediction problem.

@@

Some significant gene expression predictors were closely related to PTEN
and the PI3K pathway. PTEN and IGFR1R are the strongest gene expression
predictor which have negative and positive predictive power. IGF1R is a
tyrosine kinase receptor that activates PI3K. {[}14{]} Insulin receptor
substrate-2 (IRS2) is the adaptor protein of IGF1R. {[}15{]} PTEN is an
important regulator of PIP\textsubscript{3} by dephosphorylating
PIP\textsubscript{3} in contrast PI3K.{[}6{]}

There is some study trying to predict genomic alterations from gene
expression data. A study performed PIK3CA mutation prediction by a
gene-expression signature which is a sum of the average of the
logarithmic gene expression. The model showed good performance AUROC
0.71 in an independent test set. {[}4,16{]} Another study predicted copy
number alterations with gene expression using a multinomial logistic
regression model with least absolute shrinkage and selection operator
(LASSO). The prediction of the 1p/19q codel was very good with an AUROC
of 0.997. The gene-level prediction was good with an AUROC 0.75.
{[}17{]} A Hidden Markov Model Approach for Prediction of Genomic
Alterations from Gene Expression Profiling.{[}18{]} A logistic
regression model was used for MYCN gene amplification in neuroblastoma.
{[}19{]}

\begin{itemize}
\tightlist
\item
  Conclusions comments on possible biological or clinical implications
  and suggestions for further research.
\end{itemize}

@@

\begin{enumerate}
\def\labelenumi{(\Alph{enumi})}
\tightlist
\item
  Left: receiver operating characteristic (ROC) curve right:
  precision-recall (PR) curve of trainset and test set. The horizontal
  green line is the PIK3CA mutation rate (0.11) (B) Correlation between
  trainset and test set of the area under the receiver operating
  characteristic curve (AUROC) and area under the precision-recall curve
  (AUPR) among cancer type. The abbreviations are explained in the S1
  Appendix. (C) Correlation between the PIK3CA mutation rate of area
  under the receiver operating characteristic curve (AUROC) and area
  under the precision-recall curve (AUPR).
\end{enumerate}

@@

\begin{itemize}
\tightlist
\item
  S1 Appendix.\\
\item
  S2 Figure.\\
\item
  S1 Table.\\
\item
  S2 Table.\\
\item
  S3 Table.
\end{itemize}

@@

\leavevmode\hypertarget{ref-samurRTCGAToolboxNewTool2014}{}%
9. Samur MK. RTCGAToolbox: A new tool for exporting TCGA Firehose data.
PLoS One. 2014;9(9):e106397.

\leavevmode\hypertarget{ref-kuhnRecipesPreprocessingTools2020}{}%
10. Kuhn M, Wickham H. Recipes: Preprocessing tools to create design
matrices. 2020.

\leavevmode\hypertarget{ref-friedmanRegularizationPathsGeneralized2010}{}%
11. Friedman J, Hastie T, Tibshirani R. Regularization Paths for
Generalized Linear Models via Coordinate Descent. J Stat Softw. 2010;33:
1--22.

\leavevmode\hypertarget{ref-saitoPrecisionRecallPlotMore2015}{}%
12. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
than the ROC Plot When Evaluating Binary Classifiers on Imbalanced
Datasets. PLOS ONE. Public Library of Science; 10: e0118432.
doi:\href{https://doi.org/10.1371/journal.pone.0118432}{10.1371/journal.pone.0118432}

\leavevmode\hypertarget{ref-kuhnTidymodelsEasilyInstall2020}{}%
13. Kuhn M, Wickham H. Tidymodels: Easily install and load the
'tidymodels' packages. 2020.

\leavevmode\hypertarget{ref-leroithInsulinlikeGrowthFactor2003}{}%
14. LeRoith D, Roberts CT. The insulin-like growth factor system and
cancer. Cancer Letters. 2003;195: 127--137.
doi:\href{https://doi.org/10.1016/S0304-3835(03)00159-9}{10.1016/S0304-3835(03)00159-9}

\leavevmode\hypertarget{ref-heInteractionInsulinReceptor1996}{}%
15. He W, Craparo A, Zhu Y, O'Neill TJ, Wang L-M, Pierce JH, et al.
Interaction of Insulin Receptor Substrate-2 (IRS-2) with the Insulin and
Insulin-like Growth Factor I Receptors EVIDENCE FOR TWO DISTINCT
PHOSPHOTYROSINE-DEPENDENT INTERACTION DOMAINS WITHIN IRS-2. J Biol Chem.
American Society for Biochemistry and Molecular Biology; 1996;271:
11641--11645.
doi:\href{https://doi.org/10.1074/jbc.271.20.11641}{10.1074/jbc.271.20.11641}

\leavevmode\hypertarget{ref-cizkovaGeneExpressionProfiling2010}{}%
16. Cizkova M, Cizeron-Clairac G, Vacher S, Susini A, Andrieu C,
Lidereau R, et al. Gene Expression Profiling Reveals New Aspects of
PIK3CA Mutation in ERalpha-Positive Breast Cancer: Major Implication of
the Wnt Signaling Pathway. PLOS ONE. Public Library of Science; 5:
e15647.
doi:\href{https://doi.org/10.1371/journal.pone.0015647}{10.1371/journal.pone.0015647}

\leavevmode\hypertarget{ref-muCNAPEMachineLearning2019}{}%
17. Mu Q, Wang J. CNAPE: A Machine Learning Method for Copy Number
Alteration Prediction from Gene Expression. IEEE/ACM Transactions on
Computational Biology and Bioinformatics. 2019; 1--1.
doi:\href{https://doi.org/10.1109/TCBB.2019.2944827}{10.1109/TCBB.2019.2944827}

\leavevmode\hypertarget{ref-gengHiddenMarkovModel2008}{}%
18. Geng H, Ali HH, Chan WC. A Hidden Markov Model Approach for
Prediction of Genomic Alterations from Gene Expression Profiling. In:
Măndoiu I, Sunderraman R, Zelikovsky A, editors. Bioinformatics Research
and Applications. Berlin, Heidelberg: Springer; 2008. pp. 414--425.
doi:\href{https://doi.org/10.1007/978-3-540-79450-9_38}{10.1007/978-3-540-79450-9\_38}

\leavevmode\hypertarget{ref-heGeneSignaturesAssociated2020}{}%
19. He X, Qin C, Zhao Y, Zou L, Zhao H, Cheng C. Gene signatures
associated with genomic aberrations predict prognosis in neuroblastoma.
Cancer Communications. 2020;40: 105--118.
doi:\href{https://doi.org/10.1002/cac2.12016}{10.1002/cac2.12016}

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Prediction of PIK3CA mutation with gene expression} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Jun Kang\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital, College of
Medicine, The Catholic University of Korea, Seoul, South Korea}},
Ahwon Lee\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital, College of
Medicine, The Catholic University of Korea, Seoul, South Korea}},
Youn Soo Lee\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital, College of
Medicine, The Catholic University of Korea, Seoul, South Korea}}\textsuperscript{*}\\
\bigskip
\bigskip
* Corresponding author: lys9908@catholic.ac.kr\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Breast cancer with PIK3CA mutation has been approved to use PIK3CA
inhibitor in hormone receptor-positive HER2 negative subtype. We apply a
supervised elastic net penalized logistic regression model in prediction
PIK3CA mutation from gene expression data. We used the TCGA pan-cancer
dataset. Penalized logistic regression was applied to prediction
modeling. 10-fold cross-validation with targe variable stratification
was done over the hyperparameter grid. 10845 cases were available for
both PIK3CA mutation and mRNA expression data. In 10-fold
cross-validation, the model with \(\lambda\) = 0.01 and \(\alpha\) = 1.0
(Ridge regression) showed best performance in terms of AUROC. The final
model was trained with the selected hyperparameters with the entire
trainset. The trainset AUROC was 0.93 and the testset AUROC was 0.84.
The AUPR of trainset was 0.66 and the testset AUPR was 0.39. Both IGFR1R
and PTEN are key players in the tyrosine kinase pathway. The cancer type
was an important predictor. Our study suggests that the prediction of
genomic alteration with gene expression data is possible with good
performance.

@@

@@

It has been actively researched that the machine learnig approach can be
applied to detect genomic alterations. Machine learning algorithm can
build a prediction model from a large number of predictors such as
radiolomic features {[}1{]}, pathology image {[}2{]} or gene expression
data {[}3{]}. Because the most direct genomic test is more specific and
sensitive than predictive models, the machine learning approach might
have a limited role in clinical practice. However, the machine learning
prediction can be a second-best when the direct test fails.

@@

We used the TCGA pan-cancer dataset. TCGA is a cancer genomic consortium
that archives data of exome sequencing, gene expression, DNA
methylation, protein expression, and clinical data of more than 10000
cancer samples across 33 common cancer types. The gene expression TCGA
pan-cancer dataset is batch-corrected with normalization. The TCGA
dataset is publically available. PIK3CA mutation data was got using
cgdsr rpackage.{[}8{]} Gene expression data was downloaded from the
National Cancer Institute (NCI)'s Genomic Data Commons (GDC) website
that archives data used for The Pan-Cancer Atlas initiative.
(https://gdc.cancer.gov/about-data/publications/pancanatlas)

@@

To narrow down potential predictors, genes with a large the median
absolute deviation (more than third-quartiles) were selected. 33 cancer
type dummy variables were included in predictor variables. Yeo-Johnson
transformation was done to correct skewness. Centering and scaling were
done. All preprocessing was done using recipe r package. {[}9{]}
Penalized logistic regression was applied to prediction modeling.
10-fold cross-validation with targe variable stratification was done
over the hyperparameter grid: \(\lambda\) \{\(10^{-5}\),
\(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
\(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Lambda is penalty scaling parameter
and alpha is mixing parameter of penalty function
(\((1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1\)).
{[}10{]}

@@

Model performance was evaluated with the area under the receiver
operating characteristic (ROC) and area under the ROC (AUROC) and
precision-recall (PR) and area under the curve (AUPR). The AUPR is more
informative than AUROC on imbalanced datasets. {[}11{]} The modeling
process and accessing model performance were done with tidymodels
rpackage. {[}12{]}

@@

\hypertarget{dataset-summary}{%
\subsection{Dataset summary}\label{dataset-summary}}

10845 cases were available for both PIK3CA mutation and mRNA expression
data. 5128 out of 20502 genes were included in the modeling process
after filtering with median absolute deviation as described at the
modeling process method. The prevalence rate of PIK3CA was 0.11 in all
cases. The PIK3CA prevalence rate of each cancer type varied. The median
prevalence rate of PIK3CA of each cancer type was 0.03 (range 0-0.33)
(Figure 1).

@@

In 10-fold cross-validation, the model with \(\lambda\) = 0.01 and
\(\alpha\) = 1.0 (Ridge regression) showed best performance in terms of
AUROC. The final model was trained with the selected hyperparameters
with the entire trainset. The trainset AUROC was 0.93 and the testset
AUROC was 0.84. The AUPR of trainset was 0.66 and the testset AUPR was
0.39. (Figure 2A)

@@

Because the prevalence of PIK3CA mutation varies across the cancer type,
the performance of each cancer type was investigated. The AUROC and AUPR
were positively correlated between the train set and test set in cancer
type subanalysis. (Figure 2B) The AUPR was high in cancer type with high
PIK3CA mutation rate such as colon, breast, uterus cancer types. The
AUROC did not correlate with the PIK3CA mutation rate of each cancer
type. (Figure 2C)

@@

Figure 3 shows the top 30 important predictors. The coefficient is the
parameter of the predictor which represents the effect of the predictor
on prediction. IGF1R mRNA expression was the strongest negative
predictor and PTEN was the strongest positive predictor. Both IGFR1R and
PTEN are key players in the tyrosine kinase pathway. The cancer type was
an important predictor. Some cancer types including uterine
carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA), pancreatic
adenocarcinoma (PAAD), lymphoid neoplasm diffuse large B-cell lymphoma
(DLBC) are the strongest predictors.

@@

Our model showed good performance to predict the PIK3CA mutation of
various cancer types. This result suggests that the supervised elastic
net penalized logistic regression model can be applied not only to the
RAS activation pathway but also other genomic alterations. Both the RAS
activation pathway and PIK3CA mutation are important and common cancer
genomic alterations. They have a significant effect on gene expression
in cancer cells. It can not be generalized that the supervised elastic
net penalized logistic regression model can be applied to other genomic
alteration. It might be a challenging prediction of genomic alterations
that are infrequent or have a weak effect on gene expression.

Prediction modeling from the TCGA pan-cancer dataset has limitations
regarding data preprocessing. Methods for cancer gene expression have
been developed for research. The gene expression data is processed
between-sample normalization to remove the batch effect. If the model
has been trained from the between-sample normalization, a new sample can
not be exactly preprocessed with normalization which was done on
trainset. A model based on gene expression TCGA pan-cancer dataset has a
limitation on data preprocessing. It is necessary for developing a
preprocessing method that is independent with a dataset to apply the
gene expression data to the prediction model.

Our PIK3CA prediction model performed similar with the RAS activation
prediction model of the previous study in terms of AUROC (0.84). However
AUPR of our model for PIK3CA is lower than model for RAS activation
(0.39 vs 0.63). The reason of lower AUPR of our model can be explained
by low prevalence rate of PIK3CA mutation and imbalanced dataset. The
model for RAS activation trained with cancer types with more than 0.05
prevalence of RAS activation to avoid imbalance classification problem.
We included all cancer types into modeling process. The lower prevalance
rate of target variable mean our dataset has a lower baseline of AUPR.
In the subanalysis for performance of each cancer types, the cancer
types with higher PIK3CA mutation rate showe better AUPR.

@@

Some significant gene expression predictors were closely related to PTEN
and the PI3K pathway. PTEN and IGFR1R are the strongest gene expression
predictor which have negative and positive predictive power. IGF1R is a
tyrosine kinase receptor that activates PI3K. {[}13{]} Insulin receptor
substrate-2 (IRS2) is the adaptor protein of IGF1R. {[}14{]} PTEN is an
important regulator of PIP\textsubscript{3} by dephosphorylating
PIP\textsubscript{3} in contrast PI3K.{[}6{]}

There is some study trying to predict genomic alterations from gene
expression data. A study performed PIK3CA mutation prediction by a
gene-expression signature which is a sum of the average of the
logarithmic gene expression. The model showed good performance AUROC
0.71 in an independent test set. {[}4,15{]} Another study predicted copy
number alterations with gene expression using a multinomial logistic
regression model with least absolute shrinkage and selection operator
(LASSO). The prediction of the 1p/19q codel was very good with an AUROC
of 0.997. The gene-level prediction was good with an AUROC 0.75.
{[}16{]} A Hidden Markov Model Approach for Prediction of Genomic
Alterations from Gene Expression Profiling.{[}17{]} A logistic
regression model was used for MYCN gene amplification in neuroblastoma.
{[}18{]}

@@

\begin{enumerate}
\def\labelenumi{(\Alph{enumi})}
\tightlist
\item
  Left: receiver operating characteristic (ROC) curve right:
  precision-recall (PR) curve of trainset and test set. The horizontal
  green line is the PIK3CA mutation rate (0.11) (B) Correlation between
  trainset and test set of the area under the receiver operating
  characteristic curve (AUROC) and area under the precision-recall curve
  (AUPR) among cancer type. The gray band is 95\% confidence interval.
  The abbreviations are explained in the S1 Appendix. (C) Correlation
  between the PIK3CA mutation rate of area under the receiver operating
  characteristic curve (AUROC) and area under the precision-recall curve
  (AUPR).
\end{enumerate}

@@

\begin{itemize}
\tightlist
\item
  S1 Appendix.
\end{itemize}

@@

\leavevmode\hypertarget{ref-kuhnRecipesPreprocessingTools2020}{}%
9. Kuhn M, Wickham H. Recipes: Preprocessing tools to create design
matrices. 2020.

\leavevmode\hypertarget{ref-friedmanRegularizationPathsGeneralized2010}{}%
10. Friedman J, Hastie T, Tibshirani R. Regularization Paths for
Generalized Linear Models via Coordinate Descent. J Stat Softw. 2010;33:
1--22.

\leavevmode\hypertarget{ref-saitoPrecisionRecallPlotMore2015}{}%
11. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
than the ROC Plot When Evaluating Binary Classifiers on Imbalanced
Datasets. PLOS ONE. Public Library of Science; 10: e0118432.
doi:\href{https://doi.org/10.1371/journal.pone.0118432}{10.1371/journal.pone.0118432}

\leavevmode\hypertarget{ref-kuhnTidymodelsEasilyInstall2020}{}%
12. Kuhn M, Wickham H. Tidymodels: Easily install and load the
'tidymodels' packages. 2020.

\leavevmode\hypertarget{ref-leroithInsulinlikeGrowthFactor2003}{}%
13. LeRoith D, Roberts CT. The insulin-like growth factor system and
cancer. Cancer Letters. 2003;195: 127--137.
doi:\href{https://doi.org/10.1016/S0304-3835(03)00159-9}{10.1016/S0304-3835(03)00159-9}

\leavevmode\hypertarget{ref-heInteractionInsulinReceptor1996}{}%
14. He W, Craparo A, Zhu Y, O'Neill TJ, Wang L-M, Pierce JH, et al.
Interaction of Insulin Receptor Substrate-2 (IRS-2) with the Insulin and
Insulin-like Growth Factor I Receptors EVIDENCE FOR TWO DISTINCT
PHOSPHOTYROSINE-DEPENDENT INTERACTION DOMAINS WITHIN IRS-2. J Biol Chem.
American Society for Biochemistry and Molecular Biology; 1996;271:
11641--11645.
doi:\href{https://doi.org/10.1074/jbc.271.20.11641}{10.1074/jbc.271.20.11641}

\leavevmode\hypertarget{ref-cizkovaGeneExpressionProfiling2010}{}%
15. Cizkova M, Cizeron-Clairac G, Vacher S, Susini A, Andrieu C,
Lidereau R, et al. Gene Expression Profiling Reveals New Aspects of
PIK3CA Mutation in ERalpha-Positive Breast Cancer: Major Implication of
the Wnt Signaling Pathway. PLOS ONE. Public Library of Science; 5:
e15647.
doi:\href{https://doi.org/10.1371/journal.pone.0015647}{10.1371/journal.pone.0015647}

\leavevmode\hypertarget{ref-muCNAPEMachineLearning2019}{}%
16. Mu Q, Wang J. CNAPE: A Machine Learning Method for Copy Number
Alteration Prediction from Gene Expression. IEEE/ACM Transactions on
Computational Biology and Bioinformatics. 2019; 1--1.
doi:\href{https://doi.org/10.1109/TCBB.2019.2944827}{10.1109/TCBB.2019.2944827}

\leavevmode\hypertarget{ref-gengHiddenMarkovModel2008}{}%
17. Geng H, Ali HH, Chan WC. A Hidden Markov Model Approach for
Prediction of Genomic Alterations from Gene Expression Profiling. In:
Măndoiu I, Sunderraman R, Zelikovsky A, editors. Bioinformatics Research
and Applications. Berlin, Heidelberg: Springer; 2008. pp. 414--425.
doi:\href{https://doi.org/10.1007/978-3-540-79450-9_38}{10.1007/978-3-540-79450-9\_38}

\leavevmode\hypertarget{ref-heGeneSignaturesAssociated2020}{}%
18. He X, Qin C, Zhao Y, Zou L, Zhao H, Cheng C. Gene signatures
associated with genomic aberrations predict prognosis in neuroblastoma.
Cancer Communications. 2020;40: 105--118.
doi:\href{https://doi.org/10.1002/cac2.12016}{10.1002/cac2.12016}
